---
title: "DATA5002 Week 7 Lab"
subtitle: "Interactive visualisation"
date: "29 October 2024"
output:
  html_notebook: default
---

# Instructions

The purpose of this lab is to practice creating and customising JavaScript-based interactive visualisations.

You will require the following packages, though others may also be helpful.

* `rmarkdown`
* `tidyverse`
* `plotly`
* `crosstalk`
* `ggiraph`

Running the following script will install them if not already installed:

```{r eval=FALSE}
required_packages <- c("rmarkdown", "tidyverse", 
                       "plotly", "crosstalk", "ggiraph")
# installed.packages() is a data frame with info about 
# versions and other things we don't need.
installed_packages <- rownames(installed.packages()) 
# Which ones are required but not installed?
to_install <- setdiff(required_packages, installed_packages)
# Install them.
install.packages(to_install) 
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr) 
library(ggplot2)
library(plotly)
```

Please work through the following exercises. As before, the "Hints" tab will contain the list of functions and elements that I would use.

# Exercises

## Part 1: Using `ggplotly()`


#### **(a)**
```{r}



```



#### **(b)**
```{r}
scoring_rate <- read.csv("scoring_rate.csv")
p <- scoring_rate %>% ggplot(aes(x=Minutes, y=Scored/Minutes))+ geom_point(size=0.3) + labs(x="Player minutes",y="scored per minute")

p
ggplotly(p,tooltip="text")

```
## Part 2: Using `plot_ly()`


#### **(a)**
```{r}


plot_ly(scoring_rate,x=~Minutes, y=~Scored/Minutes,type="scatter")

plot_ly(scoring_rate,x=~Minutes, y=~Scored/Minutes,type="scatter") %>% add_markers( )
for (i in 1:nrow(scoring_rate)){
if (scoring_rate[i,1]>=2010) {  
scoring_rate[i,5] <- 2010 }else {scoring_rate[i,5] <- 2000}
  i=i+1
}

# To make code more readable, plotly imports the pipe operator from magrittr
economics %>% plot_ly(x = ~date, y = ~unemploy/pop) %>% add_lines()

# Attributes defined via plot_ly() set 'global' attributes that 
# are carried onto subsequent traces, but those may be over-written
plot_ly(economics, x = ~date, color = I("black")) %>%
 add_lines(y = ~uempmed) %>%
 add_lines(y = ~psavert, color = I("red"))



```




#### **(b)**

```{r}



```


## Part 3: Dashboard


#### **(a)**

```{r}



```



### Part 4: Crosstalk

```{r}



```

#### **(a)**

```{r}



```


## Part 5: ggiraph

Note: some of the more powerful features of `ggiraph` work best when incorporated into a shiny app.
```{r}



```


#### **(a)**

```{r}



```




## Part 6: Combining plots (Optional)

```{r}



```

#### **(a)**

```{r}



```



#### **(b)**

```{r}



```



